---
title: "PhilaFire_PropensityScore"
author: "Ben Keel"
date: "2023-03-20"
output: html_document
---

# Set Up

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(boxr)
library(mapview)
library(sf)
library(tidyverse)
library(knitr)
library(kableExtra)
library(tigris)
library(viridis)
library(dplyr)
library(tidycensus)
library(ggplot2)
library(RSocrata)
library(lubridate)
library(janitor)
library(MatchIt)

options(scipen = 999)
mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))

#Color Palettes
palette3_sat <- c("#e19825","#d55816","#7b230b")
palette3_desat <- c("#B19C7D","#7F5F52","#262626")

palette4 <- c("#f1c82b","#e19825","#d55816","#7b230b")
palette4_desat <- c("#B19C7D","#B27D49","#7F5F52","#262626")

palette5_sat <- c("#f1c82b","#e19825","#d55816","#7b230b","#413028")
pallette5_desat <- c("#ead5b7","#d2b190","#b18e6f","#7f5f52","#413028")

palette6_sat <- c("white","#f1c82b","#e19825","#d55816","#7b230b","#413028")


palette7_cats <- c("#b9cfcf","#20b1ae","#e19825","#7b230b","#b47c49", "#3f3128", "#8f8172")

#Sources for Graphs
creditFire <- "Source: Philadelphia Fire Department"
creditOpen <- "Source: Open Data Philly"

g<-glimpse

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

crs <- "EPSG:4326"
PHL_crs <- 'ESRI:102411' #This measurement-based CRS projection is for the West Coast

```

#Process

1. Load in results, as we captured in EDA.

2. Add open data covariates (OPA, ACS, Transfer Values) to those results.

3. Check mean (0 to 1) or of results for each test group. 

# Tying the Sample Code to Our Own Data

<https://sejdemyr.github.io/r-tutorials/statistics/tutorial8.html>

## In the sample data:

Outcome variable of interest: standardized math score

Independent variable of interest: Catholic School y/n

Covariates:

-   race_white: Is the student white (1) or not (0)?

-   p5hmage: Mother's age

-   w3income: Family income

-   p5numpla: Number of places the student has lived for at least 4 months

-   w3momed_hsb: Is the mother's education level high-school or below (1) or some college or more (0)?

## In our data:

Outcome variable of interest: Vacancy / Permits / Transfers

Independent variable of interest: Fire y/n

-   Structure Fires Spatialized

    -   address (ID)

Covariates:

-   OPA Data (Tab Join)

    -   address (ID)
    
    -   building_code (categorical)
    
    -   total_livable_area (continuous)

    -   number_of_bedrooms (continuous)
    
    -   number_of_bathrooms (continuous)
    
    -   year_built (continuous)
    
    -   quality_grade (ordinal)
    
    -   owner occupied (t/f)
    
    -   assessment price/sqft (continuous)

-   ACS Data (Spatial Join)

    -   Tract Average Income (continuous)
    
    -   Tract average White racial makeup % (continuous)
    
-   Neighborhoods (Spatial Join)

    -   NHood Name
    
-   Transfers Data (optional)
    
    -   Sales price of 5 nearest buildings (continous) 

#Data Imports

## Fire Data

### Loading In

```{r Box Set Up, message=FALSE, warning=FALSE}

```

### Cleaning

```{r Cleaning The Dataset}
# Creating geometry for the fires
dat <- structureFire

dat <- dat %>% drop_na("Longitude", "Latitude") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = "EPSG:4326")

# Cleaning the column names
dat <- clean_names(dat)

# Address string
dat <-
  dat %>% mutate(street_type = ifelse(street_type == 'AV', "AVE", street_type)) %>% 
  unite(address, c('address_number', 'street_prefix', 'street_name', 'street_type'), sep = " ", remove = FALSE, na.rm=TRUE)

# Extracting quarter
dat <- dat %>% mutate(quarter = floor_date(alarm_date, unit="quarter"))

# Reducing columns
dat <- dat %>%
  dplyr::select(address, quarter, property_use, incident_number, number_of_exposures, incident_type, building_status, fire_spread, no_spread, code_description, geometry, alarm_date, cad_nature_code_description,
                minor_damage, significant_damage, heavy_damage, extreme_damage)

# Removing duplicates
dat <- dat[!duplicated(dat$incident_number),]

dat_boolean <- dat%>%
  st_drop_geometry%>%
  mutate(fire = 1)%>%
  dplyr::select(address, fire)

taney <- dat %>% filter(grepl("N TANEY", address))

rowhomes <- opa_ps %>% filter(grepl("ROW", building_code_description))

nrow(rowhomes)/nrow(opa_dat)

```

## ACS Philadelphia

### Variables

```{r Key and Variables}

census_api_key("3c9540be1434ac4b38e6e55d60e8ee95909f2254", overwrite = TRUE)

#Variables
acs_variable_list.2020 <- load_variables(2020, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

acs_variable_list.2016 <- load_variables(2016, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

```

### Loading

```{r ACS Data Loading, warning=FALSE, message=FALSE}

acs_vars <- c("B02001_001E", #Total Population
              "B02001_002E", #Total Population White Only
              "B19013_001E", #Median HH Income
              "B25002_001E", #Total Units
              "B25002_003E" #Total Vacant Units
              )

acsTractsPHL.2020 <- get_acs(geography = "tract",
                             year = 2020, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide")%>%
  rename(Total_Pop = B02001_001E,
         Total_Pop_WhiteOnly = B02001_002E,
         Med_Income = B19013_001E,
         Total_Vacant = B25002_003E,
         Total_Units = B25002_001E)%>%
  mutate(Perc_White = Total_Pop_WhiteOnly/Total_Pop,
         Perc_Vacant = Total_Vacant*100/Total_Units)%>%
  dplyr::select(-B02001_001M, -B02001_002M, -B19013_001M, -Total_Pop_WhiteOnly)%>%
  st_set_crs("EPSG:4326")

```

## OPA Data

### Loading

```{r OPA Data Import}
# Reading the data
opa_dat <- read_csv("../Data/OpenDataPhilly-opa_properties_public.csv")

# Creating geometry for the properties
opa_dat <- opa_dat%>%
  drop_na(lat, lng)%>%
  st_as_sf(coords = c("lat", "lng"),
           crs = "EPSG:4326")

```

### Cleaning

```{r OPA Data Cleaning}

# Reducing columns
opa_ps <- opa_dat[!duplicated(opa_dat$location),] %>%
  dplyr::select(location, category_code, category_code_description, building_code, building_code_description, total_area, total_livable_area, market_value, mailing_street, number_of_bedrooms, number_of_bathrooms, number_stories, year_built, quality_grade)%>%
  filter(category_code == 1 | category_code == 2 | category_code == 3)%>%
  mutate(Price_Sqft = ifelse(is.na(total_livable_area) == FALSE & total_livable_area != 0, market_value / total_area, NA),
         quality_grade_mod = case_when(quality_grade == 1 ~ "E",
                                   quality_grade == 2 ~ "D",
                                   quality_grade == 3 ~ "C",
                                   quality_grade == 4 ~ "B",
                                   quality_grade == 5 ~ "A",
                                   quality_grade == 6 ~ "A+", 
                                   TRUE ~ quality_grade,
                                   ))%>%
  rename(address = location)%>%
  mutate(condo = ifelse(grepl("CONDO", building_code_description) == TRUE, TRUE, FALSE),
        owner_occ = ifelse(condo == FALSE & category_code_description != "MULTI FAMILY",
                           ifelse(address == mailing_street, TRUE, FALSE),
                           NA))%>%
  filter(Price_Sqft < 100000)%>%
  dplyr::select(-quality_grade)

opa_ps$quality_grade_mod <- factor(opa_ps$quality_grade_mod, order = TRUE,
                               levels = c("A+", "A", "A-", "B+", "B", "B-", "C+", "C", "C-", "D+", "D", "D-", "E+", "E", "E-"))


```

## Neighborhoods

### Loading In

```{r}

#https://github.com/azavea/geo-data

nhoods <- st_read("../Data/Neighborhoods_Philadelphia.geojson")%>%
  st_set_crs("EPSG:4326")

```


## Combining Data

```{r Tabular and Spatial Join }

#Fires
opa_fire <- opa_ps%>%
  left_join(dat_boolean, by="address")

opa_fire$fire <- as.factor(ifelse(is.na(opa_fire$fire), 0, 1))

#ACS and NHoods
opa_sf <- opa_fire%>%
  st_join(acsTractsPHL.2020)%>%
  st_join(nhoods)

```

# PreTests
```{r }

g(filter(opa_sf, total_livable_area > 10000))

opa_sf_cov <- c('total_livable_area', 'number_of_bedrooms', 'Price_Sqft', 'Med_Income', 'owner_occ', 'Perc_White', 'quality_grade_mod', 'name')

opa_sf %>%
  filter(total_livable_area < 10000,
       Price_Sqft < 10000)%>%
  st_drop_geometry()%>%
  group_by(fire) %>%
  dplyr::select(one_of(opa_sf_cov)) %>%
  summarise_all(funs(mean(., na.rm = T)))

```

# Propensity Score Estimation

```{r}

g(opa_sf)

m_ps <- glm(fire ~ total_livable_area + number_of_bedrooms + Price_Sqft + Med_Income + Perc_White, 
            family = binomial(), data = opa_sf)

summary(m_ps)


m_ps2 <- glm(fire ~ total_livable_area + number_of_bedrooms + Price_Sqft + Med_Income + Perc_White + owner_occ + quality_grade_mod + name, 
            family = binomial(), data = opa_sf)

summary(m_ps2)

```

```{r}

prs_df <- data.frame(pr_score = predict(m_ps2, type = "response"),
                     fire = m_ps2$model$fire)
head(prs_df)

```

# REgion of common support

```{r}

labs <- paste("Actual school type attended:", c("Fire", "No Fire"))
prs_df %>%
  mutate(fire = ifelse(fire == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score)) +
  geom_histogram(color = "white") +
  facet_wrap(~fire) +
  xlab("Probability of Having A Fire") +
  theme_bw()

```

# Matching Algorithm

```{r}

opa_nomiss <- opa_sf %>%  # MatchIt does not allow missing values
  dplyr::select(address, fire, one_of(opa_sf_cov)) %>%
  na.omit()

mod_match <- matchit(fire ~ total_livable_area + number_of_bedrooms + Price_Sqft + Med_Income + Perc_White + owner_occ + quality_grade_mod + name,
                     method = "nearest", data = opa_nomiss)

```

```{r}

dta_m <- match.data(mod_match)

```

#Difference In Means

```{r}

dta_m %>%
  st_drop_geometry()%>%
  group_by(fire) %>%
  dplyr::select(one_of(opa_sf_cov)) %>%
  summarize_all(funs(mean))

```

# Calculating Differences Between Prop-Matched Structures

## Loading In Data

```{r Loading Panel}

panel_opa <- read_csv("../Data/panel_OPA_Fire_OpenData.csv")

```
## Filter Panel By Addresses in Prop-Matched Set

```{r}

#assign fire time and incident number to one member of the subclass
dta_m_1 <- dta_m %>%
  filter(fire == 1)%>%
  left_join(st_drop_geometry(dplyr::select(dat, incident_number, address, quarter)))

#assign same data to other member of the subclass
dta_m_2 <- dta_m %>%
  filter(fire == 0)%>%
  left_join(st_drop_geometry(dplyr::select(dta_m_1, incident_number, quarter, subclass)), by="subclass")

dta_m_withTime <- rbind(dta_m_1, dta_m_2) #combine sets

panel_pm_4Y <- panel_opa%>%
  inner_join(dta_m_withTime, by="address")%>%
  mutate(mSinceFire = interval(quarter.y, quarter.x) %/% months(1),
         ySinceFire = mSinceFire / 12)%>%
  filter(mSinceFire >= -49,#Eliminate entries before fires (occurs because of incident_number group duplicates)
         mSinceFire < 49, #Eliminate entries after four years, as they are irrelevant (arbitrary)
         !(count > 0 & ySinceFire > 0))%>% #For addresses with multiple incidents, take out repeated fire observ's
  st_as_sf()

panel_pm_4Y$count <- as.factor(panel_pm_4Y$count)


panel_pm_4YForward <- panel_opa%>%
  inner_join(dta_m_withTime, by="address")%>%
  mutate(mSinceFire = interval(quarter.y, quarter.x) %/% months(1),
         ySinceFire = mSinceFire / 12)%>%
  filter(mSinceFire >= -1,#Eliminate entries before fires (occurs because of incident_number group duplicates)
         mSinceFire < 49, #Eliminate entries after four years, as they are irrelevant (arbitrary)
         !(count > 0 & ySinceFire > 0))%>% #For addresses with multiple incidents, take out repeated fire observ's
  st_as_sf()


#2 Year Outcomes for Each Incident
panel_pm_Results2Y <- panel_pm_4YForward %>%
    st_drop_geometry()%>%
    filter(., ySinceFire <= 2)%>%
    group_by(address, incident_number, subclass, fire)%>%
    summarize(count = ifelse(sum(count) > 0, 1, 0),
              severity_index = max(severity_index),
              outcome_vacant = sum(n_Vacant),
              outcome_permit = sum(n_permits),
              outcome_transfer = sum(n_transfers),
              quarter = min(quarter.x))
  

```

#Graph Outcome Proximity to Fires

```{r}

#Distribution of Outcomes, Measuring Time between Fire and Outcome
panel_pm_4Y%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  st_drop_geometry%>%
  ungroup()%>%
  filter(!(n_Vacant == 0 & n_permits == 0 & n_transfers == 0))%>%
  dplyr::select(n_transfers, n_Vacant, n_permits, ySinceFire, fire)%>%
  gather(Variable, value, -ySinceFire, -fire)%>%
  filter(value > 0)%>%
    count(ySinceFire, fire, Variable)%>%
  ggplot(aes(x=ySinceFire, y=n, fill=fire))+
    geom_col(position="dodge", just = 0.5)+
      labs(title="Permits, Vacancies, and Sales By Time From Fire, Propensity-Matched Set",
       subtitle="Philadelphia County, 2013-2018, By Quarter")+
      geom_smooth(se=FALSE, aes(color = fire, group = paste(fire, (ySinceFire >= 0))), size = 1.5)+
  facet_wrap(~Variable) + 
  xlab("Years Since Fire")+
  ylab("Count of Outcomes")+
  scale_x_continuous(breaks = c(-4, -3, -2, -1, 0, 1, 2, 3, 4))+
    theme(panel.background = element_rect(fill = "#f3efe0"))+
  scale_fill_manual(values = c("#b18e6f", "#b9cfcf"))+
  scale_color_manual(values = c("#3f3128", "#20b1ae"))+
  geom_vline(xintercept = 0, color = "#d55816", size = 1, linetype = "dashed")


```
## Map of Same Outcomes (Forward)

```{r}

#Fires
panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  filter(fire == 1)%>%
  ggplot()+
    geom_sf(data=acsTractsPHL.2020, fill='#f3efe0', color='dark gray')+
    geom_sf(color = "#20b1ae", size=0.9, alpha = 0.5)+
    labs(title = "Fires with Property Transfers, by Years since Fire") +
    mapTheme()

#No Fires
panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  filter(fire == 0)%>%
  ggplot()+
    geom_sf(data=acsTractsPHL.2020, fill='#f3efe0', color='dark gray')+
    geom_sf(color = "#3f3128", size=0.9, alpha = 0.5)+
    labs(title = "Fires with Property Transfers, by Years since Fire") +
    mapTheme()

#Distribution of Outcomes, Measuring Time between Fire and Outcome
panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012,
         fire == 1)%>%
  filter(!(n_Vacant == 0 & n_permits == 0 & n_transfers == 0))%>%
  dplyr::select(n_transfers, n_Vacant, n_permits, ySinceFire, fire)%>%
  gather(Variable, value, -ySinceFire, -fire, -geometry)%>%
  filter(value > 0)%>%
  ggplot()+
    geom_sf(data=acsTractsPHL.2020, fill='#f3efe0', color='dark gray')+
    geom_sf(aes(color=ySinceFire), size=0.9)+
    scale_color_viridis_c()+
    facet_wrap(~Variable)+
    labs(title = "Fires with Property Transfers, by Years since Fire") +
    mapTheme()
  
#Distribution of Outcomes, Measuring Time between Fire and Outcome
panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012,
         fire == 0)%>%
  filter(!(n_Vacant == 0 & n_permits == 0 & n_transfers == 0))%>%
  dplyr::select(n_transfers, n_Vacant, n_permits, ySinceFire, fire)%>%
  gather(Variable, value, -ySinceFire, -fire, -geometry)%>%
  filter(value > 0)%>%
  ggplot()+
    geom_sf(data=acsTractsPHL.2020, fill='#f3efe0', color='dark gray')+
    geom_sf(aes(color=ySinceFire), size=0.9)+
    scale_color_viridis_c()+
    facet_wrap(~Variable)+
    labs(title = "Prop-Matched Set, No Fires, by Years since Fire") +
    mapTheme()
  

```



```{r Vacancy Clusters of reporting time}

#Filter results between 0 and 1 year, 1.25 and 2 years, 2.25 and 3 years, 3.25 and 4 years

panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  filter(!(n_Vacant == 0 & n_permits == 0 & n_transfers == 0))%>%
  mutate(ySinceFire_floor = floor(ySinceFire))%>%
  dplyr::select(n_transfers, n_Vacant, n_permits, fire, ySinceFire_floor)%>%
  gather(Variable, value, -ySinceFire_floor, -fire, -geometry)%>%
  filter(value > 0)%>%
  ggplot()+
    geom_sf(data=acsTractsPHL.2020, fill='#f3efe0', color='dark gray')+
    geom_sf(aes(color=fire), size=0.75)+
    scale_color_manual(values= c("#3f3128", "#e19825"))+
    facet_wrap(~Variable + ySinceFire_floor, nrow=3, ncol=5)+
    labs(title = "Permits in Properties With Fires, by Years After The Fire") +
    mapTheme()

```

## Where does recovery happen fastest or slowest?

We see that with both vacancies and permits, the falling likelihood of reports between 6 and 18 months levels off to a sustained, slowly decreasing amount. Perhaps these are different trends, in that permits and vacancies that happen within the first year are a natural part of fire recovery, but reports after that point indicate there are obstacles to recovery. [One fire repair service](https://jenkinsrestorations.com/7-things-to-do-after-a-house-fire/#toggle-id-4-closed) lists "large fire" recovery as taking "a significant amount of time, typically at minimum several weeks, with the likelihood of it taking several months for repairs." [Others](https://exactrecon.com/how-long-does-it-take-to-repair-a-house-after-a-fire/) have similar timelines. 

Knowing that that permits have been delayed from the usual recovery timeline, we were curious if there were spatial or demographic differences among repair speed. 

We mapped these permit reports across Philadelphia and separated the data by whether it occurred within a year or between 1-4 years after the fire. 

```{r Vacancy Clusters of reporting time}

palette5_permits <- c("#BAE4B6", "#85C07F", "#5E8C59", "#4A7246", "#2D4A2A")

#Divide permits by before 1 and after 1 year

delayed.labs <- c("0-1 Years", "1-4 Years")
names(delayed.labs) <- c(0, 1)

panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  filter(!(n_permits == 0))%>%
  mutate(delayed = ifelse(ySinceFire <= 1, 0, 1))%>%
  dplyr::select(n_permits, fire, delayed)%>%
  gather(Variable, value, -delayed, -fire, -geometry)%>%
  filter(value > 0)%>%
  ggplot()+
    geom_sf(data=acsTractsPHL.2020, aes(fill=q5(Med_Income)), color=NA)+
    scale_fill_manual(values=palette5_permits, labels=qBr(acsTractsPHL.2020, "Med_Income"))+
    geom_sf(aes(color=fire), size=0.75)+
    scale_color_manual(values= c("#20B1AE", "#F1C82B"))+
    facet_wrap(~delayed, nrow=1, labeller = labeller(delayed = delayed.labs))+
    labs(title = "Permits in Properties With Fires, by Years After The Fire") +
    mapTheme()



```
There isn't an obvious spatial with  trend other than a slight decrease of repairs in Northeast and the Northern South Philadelphia.

On the vacancy end, fire stricken properties appear to have more vacancy reports in areas of already-high vacancy rates, as defined by ACS Data for 2020. Vacant properties tend to linger in these areas as well. The two maps below divide the vacancy reports for the matched-set between reports within 1 year of the fire and reports between 1 year and 4 years after the fire. The blue dots appear to somewhat consolidate into those darker, higher-vacancy areas.  

```{r vacancy clusters}

palette5_vacant <- c("#F0EBDE", "#F9F8EB", "#F0EBDE", "#D1BCA6", "#3A352F")

#Divide vacancies by before 1 and after 1 year

lingering <- c("0-1 Years", "1-4 Years")
names(lingering) <- c(0, 1)

panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  filter(!(n_Vacant == 0))%>%
  mutate(lingering = ifelse(ySinceFire <= 1, 0, 1))%>%
  dplyr::select(n_Vacant, fire, lingering)%>%
  gather(Variable, value, -lingering, -fire, -geometry)%>%
  filter(value > 0)%>%
  ggplot()+
    geom_sf(data=acsTractsPHL.2020, aes(fill=q5(Perc_Vacant)), color=NA)+
        scale_fill_manual(values=palette5_vacant, labels=qBr(acsTractsPHL.2020, "Perc_Vacant"))+
    geom_sf(aes(color=fire), size=0.75)+
    scale_color_manual(values= c("#20B1AE", "#F1C82B"))+
    facet_wrap(~lingering, nrow=1, labeller = labeller(lingering= lingering))+
    labs(title = "Matched-Set Vacancy Reports Cluster in Areas of High Vacancy",
         subtitle = "Philadelphia County, Fires from 2013-2018, Vacancy Reports from 2013-2022",
         credit = "Source: US ACS and Philadelphia Fire Department") +
    mapTheme()

```


To be more precise, we measured each neighborhood's rate of its properties'permit requests and vacancy records 1-4 years after a fire and compared it to the total total count of each for the fire-stricken properties:

```{r Combo Rate Graphs 1 Year Plus Vacancies and Permits}

library(gridExtra)

delayedMap <- panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  filter(!(n_permits == 0), fire == 1)%>%
  mutate(delayed = ifelse(ySinceFire <= 1, 0, 1))%>%
  dplyr::select(n_permits, delayed, name)%>%
  st_drop_geometry()%>%
  group_by(name) %>%
  summarize(delayRate = sum(delayed)/sum(n_permits>0), 
            permits = sum(n_permits))%>%
  arrange(desc(delayRate))%>%
  filter(permits > 1)%>%
  left_join(nhoods, by="name")%>%
  st_as_sf()%>%
  ggplot()+
    geom_sf(data=nhoods, fill='#f3efe0', color='dark gray')+
    geom_sf(aes(fill=delayRate))+
    scale_fill_viridis_c()+
    labs(title = "Rate of Delayed Permits by Neighborhood",
         subtitle = "Neighborhoods with Permits Over 1 Year After A Fire vs Total Fire-Property Permits") +mapTheme()

lingeringMap <- panel_pm_4YForward%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  filter(!(n_Vacant == 0), fire == 1)%>%
  mutate(lingering = ifelse(ySinceFire <= 1, 0, 1))%>%
  dplyr::select(n_Vacant, lingering, name)%>%
  st_drop_geometry()%>%
  group_by(name) %>%
  summarize(lingeringRate = sum(lingering)/sum(n_Vacant>0), 
            vacancies = sum(n_Vacant))%>%
  arrange(desc(lingeringRate))%>%
  filter(vacancies > 1)%>%
  left_join(nhoods, by="name")%>%
  st_as_sf()%>%
  ggplot()+
    geom_sf(data=nhoods, fill='#f3efe0', color='dark gray')+
    geom_sf(aes(fill=lingeringRate))+
    scale_fill_viridis_c()+
    labs(title = "Rate of Lingering Vacancies by Neighborhood",
         subtitle = "Neighborhoods with Vacancies 1 Year After A Fire vs Total Vacancies") +
    mapTheme()

grid.arrange(delayedMap, lingeringMap, nrow=1)

```

This map shows which areas have the highest rate of vacancies that persist over a year after the fire, which signals obstacles to recovery. The map is incomplete due to the low rate of vacancies in some areas and the fact that we excluded some building types from the analysis that are clustered in Center City.


```{r lingering vacancies}


#g(panel_pm_4YForward)

```

This analysis does show some difference in these two metrics. Recovery may happen slowest in neighborhoods with the highest lingering vacancies and delayed permits. 

## Are investors buying fire-stricken properties at a higher rate than normal properties?

With our case study featuring two examples of investor-purchased homes, we were curious if this was a larger trend. In short, the transfer of properties between investors and homeowners is a minority of property transfers as a whole, but fires do affect sales in these categories in two different ways. 

```{r Importing sale transfer data}

transfers <- read_csv("../Data/OpenDataPhilly-transfers.csv")

#Select relevant fields and filter to fire data range
transfers_data <- transfers%>%
  dplyr:: select(objectid, recording_date, street_address, document_type, grantors, grantees)%>%
  filter(year(recording_date) > 2008,
         !is.na(street_address))%>%
  rename(address = street_address)%>%
  mutate(quarter = as_date(floor_date(recording_date, unit="quarter")))

```

We're defining investors as those who own above 5 properties in the OPA dataset and those that have 'LLC', 'LP', 'REO', 'Investments', 'Homes', 'Trust', 'Corp', 'Inc' in their name. From this criteria, we created a list of investors to find in the matched set.

```{r Creating an Investor List}

toMatch <- c('LLC', 'LP', 'REO', 'Investments', 'Homes', 'Trust', 'Corp', 'Inc')

investors <- unique(grep(paste(toMatch,collapse="|"), 
                        transfers_data$grantees, value=TRUE))

#Check for owners with 5+ properties
investors_opa <- opa_dat%>%
  st_drop_geometry()%>%
  dplyr::select(owner_1)%>%
  group_by(owner_1)%>%
  filter(n()>5)

investors_list_opa <- unique(investors_opa)

transfers_data <- transfers_data%>%
  mutate(investorBuy = ifelse(grantees %in% investors | grantees %in% investors_list_opa, TRUE, FALSE),
         investorSell = ifelse(grantors %in% investors | grantors %in% investors_list_opa, TRUE, FALSE))%>%
  mutate(homeownerToInvestor = ifelse(investorBuy == TRUE & investorSell== FALSE, TRUE, FALSE),
         investorToHomeowner = ifelse(investorSell == TRUE & investorBuy== FALSE, TRUE, FALSE))

#join transfers data to property by address and quarter

panel_pm_4Y_sales <- panel_pm_4Y%>%
  rename(quarter = quarter.x)%>%
  left_join(dplyr::select(transfers_data, address, quarter, homeownerToInvestor, investorToHomeowner), by=c("address", "quarter"))


```

### Homeowner to investor, investor to homeowner

Here’s the same type of graph as the property sales, but counting sales that we classify as “homeowner-to-investor” based on their name and how many properties they own. The rate of homeowner-to-investor sales is steadily increasing over time, but fires are associated with a nearly 100% increase in the first year, with the trend maintaining over time.

```{r homeowners to investors}

panel_pm_4Y_sales%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  st_drop_geometry%>%
  ungroup()%>%
  filter(!(n_transfers == 0))%>%
  dplyr::select(ySinceFire, fire, homeownerToInvestor)%>%
  gather(Variable, value, -ySinceFire, -fire)%>%
  filter(value == TRUE)%>%
    count(ySinceFire, fire, Variable)%>%
  ggplot(aes(x=ySinceFire, y=n, fill=fire))+
    geom_col(position="dodge", just = 0.5)+
      labs(title="Property Transfers: Sales from Homeowners To Investors by Time to Fire",
       subtitle="Philadelphia County, Fires in 2013-2018, Sales in 2009-2022, By Quarter")+
      geom_smooth(se=FALSE, aes(color = fire, group = paste(fire, (ySinceFire >= 0))), size = 1.5)+
  facet_wrap(~Variable) + 
  xlab("Years Since Fire")+
  ylab("Count of Outcomes")+
  scale_x_continuous(breaks = c(-4, -3, -2, -1, 0, 1, 2, 3, 4))+
    theme(panel.background = element_rect(fill = "#f3efe0"))+
  scale_fill_manual(values = c("#b18e6f", "#b9cfcf"))+
  scale_color_manual(values = c("#3f3128", "#20b1ae"))+
  geom_vline(xintercept = 0, color = "#d55816", size = 1, linetype = "dashed")

```

Conversely, if we look at investor-to-homeowner sales, there’s a lag time where investor-owned properties with fires are sold at a lower rate, but then investor sales pick up after six months. This may indicate a flip-sale trend, with a purchase and renovation period in the couple quarters, followed by increased sales from the investors. 

```{r investors to homeowners}

panel_pm_4Y_sales%>%
  filter(year(quarter.y) < 2019 & year(quarter.y) > 2012)%>%
  st_drop_geometry%>%
  ungroup()%>%
  filter(!(n_transfers == 0))%>%
  dplyr::select(ySinceFire, fire, investorToHomeowner)%>%
  gather(Variable, value, -ySinceFire, -fire)%>%
  filter(value == TRUE)%>%
    count(ySinceFire, fire, Variable)%>%
  ggplot(aes(x=ySinceFire, y=n, fill=fire))+
    geom_col(position="dodge", just = 0.5)+
      labs(title="Property Transfers: Sales from Investors To Homeowners by Time to Fire",
       subtitle="Philadelphia County, Fires in 2013-2018, Sales in 2009-2022, By Quarter")+
      geom_smooth(se=FALSE, aes(color = fire, group = paste(fire, (ySinceFire >= 0))), size = 1.5)+
  facet_wrap(~Variable) + 
  xlab("Years Since Fire")+
  ylab("Count of Outcomes")+
  scale_x_continuous(breaks = c(-4, -3, -2, -1, 0, 1, 2, 3, 4))+
    theme(panel.background = element_rect(fill = "#f3efe0"))+
  scale_fill_manual(values = c("#b18e6f", "#b9cfcf"))+
  scale_color_manual(values = c("#3f3128", "#20b1ae"))+
  geom_vline(xintercept = 0, color = "#d55816", size = 1, linetype = "dashed")

```





