---
title: "311 Edit File"
author: "Ben Keel"
date: "2023-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)

```

## Libraries & Set Up

```{r libraries & loading data, warning=FALSE, message=FALSE}

library(tidyverse)
library(tidycensus)
library(kableExtra)
library(caret)
library(knitr) 
library(pscl)
library(plotROC)
library(pROC)
library(lubridate)
library(broom)
library(stargazer)
library(ggplot2)
library(gridExtra)
library(janitor)
library(sf)
library(viridis)
library(ggthemes)

library(gganimate)

library(transformr)
library(gifski)
library(rmapshaper)

library(boxr)
library(lubridate)

options(scipen=999)
options(tigris_class = "sf")
 
g<-glimpse

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")


crs <- "EPSG:4326"
PHL_crs <- 'ESRI:102411' #This measurement-based CRS projection is for the West Coast


census_api_key("3c9540be1434ac4b38e6e55d60e8ee95909f2254", overwrite = TRUE)

palette3_sat <- c("#e19825","#d55816","#7b230b")
palette3_desat <- c("#B19C7D","#7F5F52","#262626")

palette4 <- c("#f1c82b","#e19825","#d55816","#7b230b")
palette4_desat <- c("#B19C7D","#B27D49","#7F5F52","#262626")

palette5_sat <- c("#f1c82b","#e19825","#d55816","#7b230b","#413028")
pallette5_desat <- c("#ead5b7","#d2b190","#b18e6f","#7f5f52","#413028")

palette7_cats <- c("#b9cfcf","#20b1ae","#e19825","#7b230b","#b47c49", "#3f3128", "#8f8172")

creditFire <- "Source: Philadelphia Fire Department"
creditOpen <- "Source: Open Data Philly"

```


### ACS geometry of Philadelphia

```{r ACS Data Loading}

acs_vars <- c("B01001_001E")

acsTractsPHL.2020 <- get_acs(geography = "tract",
                             year = 2020, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide")%>%
  st_set_crs("EPSG:4326")

  

```


## Loading Data

Loading data from box:

```{r loading box data, echo=FALSE}

#Box Upload Philadelphia Structure Fire
#Box Upload Philadelphia Structure Fire
box_auth(client_id = "7olfapklfw3aflgpeb2kopxj1jceogg0", 
         client_secret = "xMUYd0SngtrtzGC92cZKon2jd6yyvM9d")
box_setwd(186732420366)

structureFire <- box_read_excel(1093000179542)

g(structureFire)

structureFire_sf <- structureFire%>%
  drop_na(Longitude, Latitude)%>%
  st_as_sf(coords = c("Longitude", "Latitude"),
           crs = "EPSG:4326")%>%
  clean_names

g(structureFire_sf)

```





```{r Coversion to address}

structureFire_sf_address <-
  structureFire_sf %>%
  mutate(street_type = ifelse(street_type == 'AV', "AVE", street_type))%>%
  unite(address, c(address_number, street_prefix, street_name, street_type), sep = " ", remove = TRUE, na.rm=TRUE)

g(structureFire_sf_address)

#Taking out duplicate incident numbers
structureFire_sf_addressU <- 
    structureFire_sf_address[!duplicated(structureFire_sf_address$incident_number),]

```

```{r Fires Count}

#Count the number of Fires per address
nFires_perAddress <- structureFire_sf_addressUC%>%
  st_drop_geometry()%>%
  count(address, sort=TRUE)%>%
  left_join(dplyr::select(structureFire_sf_address, address), by="address", na.rm=TRUE)%>%
  st_as_sf()

#remove duplicates from above
nFires_perAddress <- nFires_perAddress[!duplicated(nFires_perAddress$address),]

#Barplot of counts, by count
nFires_perAddress%>%
  filter(n < 9)%>%
  ggplot()+
  geom_bar(mapping=aes(x=as.factor(n)), fill='gray')+
  labs(title="Number of Fires Per Address",
       subtitle="Philadelphia County, 2009-2022")+
  xlab("Count of Fires")+
  ylab("Number of Structures")+
  theme(panel.background = element_blank())

#Chart of Fire Counts per address
nFires_perAddress%>%
  st_drop_geometry()%>%
    gather(Variable, value, -n) %>%
    group_by(n)%>%
    summarize(NumberOfFires = n())%>%
  kable()%>%
  kable_styling()


  
```
### Counts Over Time
```{r}

sFire_sf_yq <- structureFire_sf_addressU %>%
  mutate(Alarm_Q = quarter(`Alarm Date`, with_year = FALSE))


#Line plot of Fires per quarter, Min to Max
sFire_sf_yq %>%
  ggplot(aes(x=as_date(floor_date(`Alarm Date`, unit="quarter")), fill=Alarm_Q)) +
      geom_bar()+
      labs(title = "Quarterly Count of Unique Fire Incidents",
           subtitle = "Philadelphia County, 2009 Q1 - 2022 Q4", 
           y = "Number of Fires")+   
    scale_x_date(name = "Year", date_breaks = "1 year")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Map Over Time
```{r}

#Map the Fires, per Year
fireMap <- ggplot() +
  geom_sf(data=acsTractsPHL.2020, fill='gray', color='dark gray')+
  geom_sf(data=structureFire_sf_addressUC, alpha =  0.5, size = 0.1, color='red')+
  labs(title = "Quarterly Count of Unique Fire Incidents",
       subtitle = "Philadelphia County, 2009 Q1 - 2022 Q4", 
       y = "Number of Fires")+
  mapTheme()

fireAnim1 <- fireMap +
  transition_time(as_date(year(`Alarm Date`)))

fireAnim1

```


### Map of Outliers

We will classify an outlier as an observation more than 3 standard deviations away from the population mean.

The mean number of fires per location is 1.115, weighted by the large amount of places where only 1 fire has occurred. The standard deviation of this fire population is 0.521. The result, 2.678, means any of the 293 locations with three or more fires will be classified as an outlier. Where are these outliers, and what are their characteristics?

```{r}

nFires_perAddress%>%
  summarize(mean = mean(n), standardDev = sd(n))%>%
  summarize(outliers = mean + 3*standardDev)
  
#List of counts
nFires_perAddress%>%
  st_drop_geometry()%>%
  kable()%>%
  kable_styling()


```



```{r Map of Outliers}

nFires_perAddress_Outliers <- filter(nFires_perAddress, n>2)

ggplot()+
  geom_sf(data=acsTractsPHL.2020, fill='gray', color='dark gray')+
  geom_sf(data=nFires_perAddress_Outliers, aes(color=q5(n)), alpha=0.5)+
    scale_color_manual(values=palletteO, labels=qBr(nFires_perAddress_Outliers, "n"))+
  labs(title = "Addresses With 3+ Fires") +
  mapTheme()

```

```{r Characteristics of Outliers}

#Select addresses with details: Building Status, Structure Type, Property Use and add them to the count
nFires_perAddress_Build <- nFires_perAddress_Outliers %>%
  left_join(st_drop_geometry(dplyr::select(structureFire_sf_addressU, `Building Status`, `Structure Type`, `Property Use`, address)), by="address")

#There are 293 
g(nFires_perAddress_Build)

#Plot the Building Types and Structure Types among properties with 3+ fires
dplyr::select(nFires_perAddress_Build, -address, -`Property Use`)%>%
  st_drop_geometry()%>%
  gather(Variable, value, -n)%>%
  count(Variable, value, value)%>%
  group_by(Variable)%>%
  ggplot(., aes(value, n))+
      geom_bar(position = "dodge", stat="identity") +
      facet_wrap(~Variable, scales="free") +
      labs(x="Category", y="Frequency",
           title = "Type of Buildings Among Outliers",
           subtitle = "Buildings with 3+ Fires, 2009-2022")

```



```{r}
#Plotting the Frequency of Fires based on the Outliers' property use

dplyr::select(nFires_perAddress_Build, -address, -`Building Status`, -`Structure Type`)%>%
  st_drop_geometry()%>%
  gather(Variable, value, -n)%>%
  count(Variable, value)%>%
  group_by(Variable)%>%
  ggplot(., aes(value, n))+
      geom_bar(position = "dodge", stat="identity") +
      facet_wrap(~Variable, scales="free") +
      labs(x="Category", y="Frequency",
           title = "Type of Buildings Among Addresses with 3+ Fires",
           subtitle = "Categorical features")+
          theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r Property Use All Fires}

#Building Use for All Fires
#Plotting the Frequency of Fires based on Property Use

nFires_perAddress_BuildAll <- nFires_perAddress %>%
  left_join(st_drop_geometry(dplyr::select(structureFire_sf_addressU, `Building Status`, `Structure Type`, `Property Use`, address)), by="address")


#Bar plot of property use counts
dplyr::select(nFires_perAddress_BuildAll, -address, -`Building Status`, -`Structure Type`)%>%
  st_drop_geometry()%>%
  gather(Variable, value, -n)%>%
  count(Variable, value)%>%
  group_by(Variable)%>%
  filter(n > 150)%>%
  ggplot(., aes(value, n))+
      geom_bar(position = "dodge", stat="identity") +
      labs(x="Category", y="Frequency",
           title = "Top 10 Property Uses Among All Structure Fires",
           subtitle = "Philadelphia County, 2009-2022")+
          theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Chart of property use counts
dplyr::select(nFires_perAddress_BuildAll, -address, -`Building Status`, -`Structure Type`)%>%
  st_drop_geometry()%>%
  gather(Variable, value, -n)%>%
  count(Variable, value, sort = TRUE)%>%
  group_by(Variable)%>%
  summarize(`Property Use` = value, Total = n)%>%
  dplyr::select(`Property Use`, Total)%>%
  kable()%>%
  kable_styling()

#Bar plot of Building Status counts
dplyr::select(nFires_perAddress_BuildAll, -address, -`Property Use`, -`Structure Type`)%>%
  st_drop_geometry()%>%
  gather(Variable, value, -n)%>%
  count(Variable, value)%>%
  group_by(Variable)%>%
  ggplot(., aes(value, n))+
      geom_bar(position = "dodge", stat="identity") +
      labs(x="Category", y="Frequency",
           title = "Building Status Among All Structure Fires",
           subtitle = "Philadelphia County, 2009-2022")+
          theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Chart of Building Status counts
dplyr::select(nFires_perAddress_BuildAll, -address, -`Property Use`, -`Structure Type`)%>%
  st_drop_geometry()%>%
  gather(Variable, value, -n)%>%
  count(Variable, value, sort = TRUE)%>%
  group_by(Variable)%>%
  summarize(`Building Status` = value, Total = n)%>%
  dplyr::select(`Building Status`, Total)%>%
  kable()%>%
  kable_styling()

```

```{r Share of Property Use Amongst Super Categories}

#Count Fires by PropUse and Add Super Category column
nFires_perAddress_PropUse <- nFires_perAddress %>%
  left_join(st_drop_geometry(dplyr::select(structureFire_sf_addressU,`Property Use`, address)), by="address")%>%
  mutate(Property_Use_SuperCat = substr(`Property Use`, 1, 1))


g(nFires_perAddress_PropUse)

#Chart of Building Status counts
dplyr::select(nFires_perAddress_PropUse, -address, -`Property Use`)%>%
  st_drop_geometry()%>%
  gather(Variable, value, -n)%>%
  count(Variable, value, sort = TRUE)%>%
  group_by(Variable)%>%
  summarize(`Property Use Supercategory` = value, `Share (%)` = round((n/sum(n)*100), 2))%>%
  kable()%>%
  kable_styling()

```


```{r Map of Severity}
#Using Type as a measure of severity

nFires_perType <-  structureFire_sf_address%>%
  st_drop_geometry()%>%
  count(`Incident Type`, sort=FALSE)



nFires_perAddress %>%
  dplyr::select()



```


```{r}

#Plotting Places with Fires
ggplot()+
  geom_sf(data=acsTractsPHL.2020, fill='gray', color='dark gray')+  
  geom_sf(data=nFires_perAddress, aes(color=q5(n)), alpha=0.5, size=0.1)+
    scale_color_manual(values=palletteO, labels=qBr(nFires_perAddress, "n"))+
  labs(title = "Fires by Quantity") +
  mapTheme()

```

### Cad Nature Code Description

``` {r CAD Nature Code Description}

#Count the unique values for CAD
nFires_CADDescr <- structureFire_sf_addressUC %>%
  st_drop_geometry%>%
  count(`CAD Nature Code Description`, sort=TRUE)

nFires_CADDescr%>%
  kable()%>%
  kable_styling()

#Barplot of counts, by count
nFires_CADDescr%>%
  filter(n>50)%>%
  ggplot()+
  geom_col(mapping=aes(x=`CAD Nature Code Description`, y=n), fill='gray')+
  labs(title="Frequency of Fire Types",
       subtitle="Philadelphia County, 2009-2022, Above 50 Unique Incidents")+
  xlab("CAD Nature Code Description")+
  ylab("Number of Fires")+
            theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r Fire Damage Counts}

#How to measure severity in detail beyond 
sFire_severity <- structureFire_sf_addressUC%>%
  st_drop_geometry()%>%
  dplyr::select(`Incident Type`, `Minor Damage`, `Significant Damage`, `Heavy Damage`, `Extreme Damage`)%>%
  mutate(Worst_Damage = ifelse(`Extreme Damage` > 0, "Extreme",
                          ifelse(`Heavy Damage` > 0, "Heavy",
                            ifelse(`Significant Damage` > 0, "Significant",
                              ifelse(`Minor Damage` > 0, "Minor", "No Record")))))%>%
  count(`Incident Type`, Worst_Damage)
  
sFire_severity

table(structureFire_sf_addressUC$`Minor Damage`)
table(structureFire_sf_addressUC$`Significant Damage`)
table(structureFire_sf_addressUC$`Heavy Damage`)
table(structureFire_sf_addressUC$`Extreme Damage`)

ggplot(sFire_severity)+
  geom_col(mapping=aes(x=as.factor(`Incident Type`), y=n, fill= as.factor(Worst_Damage)))+
  labs(title = "Number of Fires by Incident Type", 
       subtitle = "Philadelphia County, 2009-2022", 
       caption = creditFire,
       x="Incident Type Code",
       y="Number of Fires")+
  scale_fill_discrete(name = "Worst Damage \nTo Floors")+
  theme(plot.title = element_text(size=18))
  



```


```{r Correlation Spread and Incident Type}

structureFire_sf_addressUC %>%
  st_drop_geometry()%>%
  count(`Incident Type`, `Fire Spread`)%>%
  filter(`Incident Type` != 111 & `Incident Type` != 1110)%>%
  ggplot()+
  geom_col(mapping=aes(x=as.factor(`Incident Type`), y=n, fill=`Fire Spread`))+
  labs(title = "Number of Fires by Incident Type, Fire Spread", 
       subtitle = "Philadelphia County, 2009-2022", 
       caption = creditFire,
       x="Incident Type Code",
       y="Number of Fires")+
  scale_fill_discrete(name = "Fire Spread \nConfined To:", 
                      labels = c("Object", "Room", "Floor", "Building", "Beyond", "NA"
  ))+
  theme(plot.title = element_text(size=18))

```
